### 2018-09-26 analysis/scripts/practice_copy.Rmd
compiled scripts that should have led to summary

---
title: "Practice"
author: "Madeleine"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#libraries
library(readr)
library(dbplyr)
library(dplyr)
library(odbc)
#path is local
#specify odbc as backend
#might also need to specify driver?
data_base <- DBI::dbConnect(odbc::odbc(), path = "~/klgo_bio_monitoring/analysis/data/raw_data")
```

```{r}
#name tables
dbListTables(data_base)
#all code like this gets blanks
test <- dbGetQuery(data_base, SELECT * FROM "Field_Data")

#creates data frame - can use tidyverse instead?
#here's an example
field_data <- dbReadTable(data_base, "Field_Data")
new_data <- field_data %>%
  group_by(Species, Breeding_Status) %>%
  summarize(count = n())
```