### 2018-12-10 analysis/paper/paper.tex
i have changed the knitting settings to put the paper in the project directory instead of the /paper folder. Now the codes for scripts run!

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{eurosym}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Assessment of the KLGO Coastal Waterbird Survey Monitoring Program},
            pdfkeywords={keyword 1; keyword 2; keyword 3},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Assessment of the KLGO Coastal Waterbird Survey Monitoring Program}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{true \\ true}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{10 December, 2018}

\begin{document}
\maketitle
\begin{abstract}
The Coastal Waterbird Survey has been conducted, with varying levels of
effort and in-season frequency, since 2003 (Surdyk and Evans
\protect\hyperlink{ref-Surdyk2018}{2018}). The area of interest has been
divided into eight contiguous survey units; on each survey, at each unit
the number, age, composition (sex), breeding status, and noteworthy
behavioral observations are recorded for each species (Anonymous
\protect\hyperlink{ref-Anonymous2009}{2009}). Given the constraints on
personnel, time, and budgets, it is imperative that surveys be conducted
efficiently in terms of both choice of sampling design and choice of
sampling effort. Now that KLGO has multiple years of experience
implementing the protocol, it is timely to step back and assess how
things are going. We review the current survey programâ\euro{}™s
objectives, design, identify statistical issues, and recommend potential
solutions. Major topics include â\euro{}¦. The study concludes with a
series of recommended tasks that the park should undertake to improve
the coastal waterbird surveyâ\euro{}™s efficiency and effectiveness.
\end{abstract}

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{Things To Discuss}\label{things-to-discuss}

\begin{itemize}
\tightlist
\item
  add all libraries used in scripts to `setup' code chunk (just above)
  rather than just loading in script so that it is easy to see what
  needs to be loaded before you can successfully knit.
\end{itemize}

\section{Introduction}\label{introduction}

Since 2003, staff at Klondike Gold Rush National Historical Park (KLGO)
have surveyed coastal waterbird species distribution and abundance
multiple times each year from late April through late September (Coastal
Waterbird Survey or `CWS'; (Surdyk and Evans
\protect\hyperlink{ref-Surdyk2018}{2018})). This is sufficient history
to warrant stepping back and assessing the monitoring program. Such
assessments are key to an adaptive monitoring program ({\textbf{???}}),
are an important and distinct phase to promote program, and
institutional, learning and evolution (Reynolds et al.
\protect\hyperlink{ref-Reynolds2016}{2016}), and align with guidance
from NPS' Inventory and Monitoring Division ({\textbf{???}}, Mitchell et
al. (\protect\hyperlink{ref-Mitchell2018}{2018})).

Assessing a monitoring program focuses on four broad topics following
the major Phases of a monitoring program (ibid):\\
1. \emph{why} - {[}Problem Framing \& Information Objectives{]} Are the
management objective(s) motivating the monitoring well articulated? Do
they remain relevant and important?\\
2. \emph{how} - {[}Data Collection Design{]} Does the monitoring data
collection design effectively addresses the information objectives?
Given interim learning regarding system characteristics and improvements
in measurement technologies, are the information goals (including
accuracy and precision goals), actually feasible? If so, can the data
collection be made more efficient, reducing the observation burden on
staff resources?\\
3. \emph{how} - {[}Data Analysis Design{]} Do the data analysis methods
effectively address the information objectives? Do the data analysis
methods properly account for impactful features of the data collection
design? Are more informative analyses available given interim
improvements in methods?\\
4. \emph{how} - {[}Workflows \& Reporting{]} Are workflows and related
software systems in place to allow timely, and efficient, data
management, analysis updating, reporting, and distribution of findings
to target audiences of information users?

This assessment addressed all four topics, but focused predominantly on
topics (i) Problem Framing \& Information Objectives, (ii) Data
Collection Design, and (iii) Data Analysis Design. Since `Problem
Framing \& Information Objectives' determine the relevancy of many of
the potential assessment actions related to the other topics, the
assessment was drafted as a first stage of an iterative effort,
providing high level feedback, including listings of potential detailed
syntheses and workflow enhancement activities (such as improved database
interface, automation of annual analyses and updating of syntheses,
etc.). With this foundation, further dialogue with KLGO staff regarding
relevant `Problem Framing \& Information Objectives' and other
priorities will determine which follow-on activities to pursue.

\section{Background}\label{background}

\section{Methods}\label{methods}

The CWS monitoring program was compared against a recommended checklist
of monitoring program components developed from published literature
(Appendix 1). It identifies multiple component decisions associated with
each of the major Phases of a monitoring program (briefly overviewed
above). For each component, the relevant aspects of the current CWS were
reviewed, potential issues identified, potentially informative
diagnostic analyses suggested, and recommendations provided. The
characterization of the CWS was based on KLGO protocols and reports
(Table 1), a site visit (6/4-6/2018) and personal communications with
the KLGO Natural Resources Program Manager (Jami Belt, winter and spring
2018), KLGO Resources Manager (Anne Matsov, summer 2018) and
long-standing survey participants (Elaine Furbish, Skagway Bird Club,
spring and summer 2018).

\section{Assessment Results}\label{assessment-results}

\subsection{Phase 1 - Problem Framing and Information
Objectives}\label{phase-1---problem-framing-and-information-objectives}

\emph{WHY} monitor? What potential decisions and issues should the
observations inform?

The management problem or question motivating the need for information
about coastal waterbirds provides the foundation for the survey's data
collection and analysis designs (Fancy, Gross, and Carter
\protect\hyperlink{ref-Fancy2009}{2009}). It clarifies the underlying
goal of the monitoring program and answers ``Why is this an important
problem or question?''.

A robust problem definition should include, among other aspects, the
temporal and geographic scope of the problem; who has the authority to
make decisions that could resolve or address the problem; what
information about the resource is needed to improve the decision making
(the information needs); and who are the stakeholders that are impacted
by the decision and/or interested in the monitoring information.
Stakeholder identification should include interested parties and their
information needs in both near-term - for example, park interpretation
and visitor education activities, as well as long-term - for example,
regional-scale changes in migration patterns or assessment and
improvement of projected climate change impacts on avifauna (e.g., Wu et
al. \protect\hyperlink{ref-Wu2018}{2018}). A good understanding of
multiple information needs and pathways to influence decision making
will guide the required data analyses and reporting and outreach.

The information needs are what the monitoring effort aims to address
(Reynolds \protect\hyperlink{ref-Reynolds2012}{2012}). Even when the
focus is simply the status and trend of a resource, there is an unstated
expectation that someone will endeavor to achieve or maintain a
desirable state or reverse an undesirable trend (Reynolds et al.
\protect\hyperlink{ref-Reynolds2016}{2016}). Whomever has that
responsibility and authority (person, office, or organization) is a
priority stakeholder. They should not only receive the reports but also
allowed an opportunity to clarify their information needs, including the
desired precision and quality of information, the timing and format of
reporting, and better understand what information is feasible to obtain
(Averill et al. 2010).

\subsubsection{Assessment findings:}\label{assessment-findings}

The 2003 coastal waterbird inventory and survey design was developed to
address ``a top priority bird inventory (need)'' of the Southeast Alaska
Inventory and Monitoring Network (Sharman et al. 2000) - namely, an
inventory of waterfowl/shorebirds at KLGO (Hahr et al.
\protect\hyperlink{ref-Hahr2004}{2004}). This was deemed a priority due
to the area's expected avifaunal richness ((NABCI)
\protect\hyperlink{ref-NABCI2000}{2000}) and the limited availability of
reliable observation records - mainly consisting of systematic surveys
conducted by NPS staff in the early 1980s (Hahr et al.
\protect\hyperlink{ref-Hahr2004}{2004}).

The inventory's stated goal was to document the occurrence of 90\% of
the species of waterbirds likely to occur in KLGO (ibid). The principal
objectives were ``to develop scientifically valid sampling designs for
inventorying waterbirds and breeding landbirds in KLGO, and to collect
baseline information on bird species occurrence, distribution,
abundance, and habitat associations in the park\ldots{}{[}to{]}
\ldots{}provide park managers with timely information that facilitates
land management planning, helps track wildlife population and ecosystem
change over time, and keeps resource managers apprised of changes that
may require action before a crisis occurs'' (ibid). While no Threatened
or Endangered bird species were known to occur in KLGO, as of 2004, nine
species on the Alaska Watchlist had been documented in the park (ibid),
thus suggesting potential priority information needs associated with
monitoring for changes in those species occurrence and abundance.
Potential secondary objectives mentioned include ``investigations into
the breeding ecology, phenology, reproductive success and population age
structure of KLGO breeding birds'' (ibid). Later reports mention
additional objectives of estimating changes in species occurrence and
composition (St. Pierre \protect\hyperlink{ref-St.Pierre2015}{2015}),
characterizing timing of peak spring waterbird migration and its
relation to spring eulachon runs (Surdyk
\protect\hyperlink{ref-Surdyk2017}{2017}).

Stakeholders mentioned or implied in park reports include park
superintendent and related managers, those charged with monitoring
regional and internationally important bird habitats, the Skagway Bird
Club, the Alaska Landbird Monitoring Survey and the Boreal Partners in
Flight.

The coastal waterbird surveys have been continued since 2003 (Surdyk and
Evans \protect\hyperlink{ref-Surdyk2018}{2018}).

QUOTES Hahr and Trapp (2004): We intend to use the results of these
inventories to create a revised bird checklist and an atlas database for
the Taiya and Skagway River watersheds in collaboration with the Skagway
Bird Club.

This protocol addresses one monitoring question with three objectives:
(1) what is the annual nest density and reproductive (fledgling) success
in known Hawaiian petrel colonies, (2) what are the long-term trends in
colony distribution and density monitored in approximate 5-year
intervals, and (3) are these affected by predator control? The first
goal of monitoring is to obtain unbiased estimates of Hawaiian petrel
nest density and reproductive (fledging) success from known colonies in
HAVO and HALE in order to detect changes in colony growth or decline.
The second goal of monitoring is to periodically (approximately every
five years) obtain unbiased estimates of nest density from potential
Hawaiian petrel habitat. This information can be used to assess changes
in density and distribution of subcolonies across the landscape. The
third goal is to estimate nest density and fledging success in areas
undergoing different management regimes to assess effectiveness of
management. Benchmark levels of these estimates could serve as warnings
of the need for modified management or further investigation of these
colonies.

Though not explicitly characterized in the protocol or reports reviewed
for this assessment (Table 1), the following target audiences and
information users were identified through conversations with KLGO staff,
partners, and other waterbird researchers and stakeholders in Alaska.
(current) - park - Skagway Bird Club; compiles for addition to
`significant bird observations in SE AK', which gets compiled in to AK
sumamries for `North American Birds' (mainly spp lists) (potential) -
BPIF - PSG - Audubon Alaska (for IDing Important Bird Areas - though
require accurate density estimates) - Audubon North America - cliamte
species distribution models (Nicole Michel, Brooke Bateman, etc.); esp
if associated w/ repeat sites \& effort data (date, time of day,
duration, distance/area covered) - see Nicole email 6/7/2018. Watchlist
of projected changes - (Wu et al. \protect\hyperlink{ref-Wu2018}{2018})

\subsubsection{Recommendations:}\label{recommendations}

Potential Follow Up Activities:

Define the \emph{Target Audiences} for the information and their
Objectives in using the information

\emph{Decision Makers} Who are target audiences for this information?
Who uses it or may use it? Local contact: C. Elaine Furbish, Skagway
Bird Club, Google Site Predominantly historical park; limited resource
staff and capacity Visible natural resource of interest to local
community and visitors; interp and educational opportunity; topic of
information requests; provides `community service' given absence of any
other agency with natural resource staff; Annual report (time lag?);
National Resource Condition Assessment; other? State of the Park?
Potential for contribution to regional, flyway, and WASO programs (e.g.,
CCRP \& Audubon climate watch, etc.) Current sightings passed to local
bird club \& used by avian-interested visitors Used in quarterly
regional `significant bird observations', which in turn is used to
develop the Alaska-wide summary for the journal North American Birds
(published by the American Birding Association
\url{http://www.aba.org/nab} ) and includes `\ldots{}sightings of birds
that are out of range or out of season or occur regularly in small
numbers, noteworthy breeding records, unusually large or small numbers
of a particular species, unusual migration dates, etc\ldots{}.' (S. C.
Heinl, 2017, Summary of SE AK Bird Observations, June-July 2017,
accessed 2018 June 6). Queries to Heather Renner \& Robin Corcoran re:
PSG-type users, etc. Opportunity for community based monitoring to track
projected spp impacts of climate change, change in phenology,
occurrence, abundance, etc.?

\emph{Decisions} What types of decisions is it / will it be used for?
(specify for each target audience/ end user) No specific decisions
Current: KLGO - spp lists, phenology of occurrence (guidance to
visitors), trends in richness (NRCA), abundance? Skagway bird club - spp
list, current occurrences (activity re: rare \& unusual) SE AK /
Regional - spp lists; unusual sightings, etc. (spp occurrences \& broad
seasonal timing). AKR - WASO/NPS - ??CCRP?? Question to Gregor \& Mel
(AK Audubon). Could be used for garnering community engagement in
Climate Watch \& other national efforts associated with Audubon's
Climate Report. Not used in continental analyses. ``One key message that
I want to emphasize is that the focus of this work was purely on climate
suitability. What this means, as we indicate in the briefs'
``Important'' box up front and ``Caveats'' section further back, is that
significant changes in climate suitability, as measured here, will not
always result in a species response, and all projections should be
interpreted as potential trends. Multiple other factors mediate
responses to climate change, including habitat availability, ecological
processes and biotic interactions, dispersal capacity, species'
evolutionary adaptive capacity, and phenotypic plasticity (e.g.,
behavioral adjustments). Ultimately, models can tell us where to focus
our concern and which species are most likely to be affected, but
monitoring is the only way to validate these projections and should
inform any on-the-ground conservation action.'' Gregor Schuurman, email
to DENA

Potential: KLGO: (easy) - changes in occurrence (spp, season scale);
changes in phenology (migration timing, by spp; breeding timing, by
spp); broad (order of magnitude) changes in abundance (index); broad (?)
changes in composition (by spp \& timing of survey)?; species occurrence
x unit (\textasciitilde{}habitat?) - refine expected lists per survey
and develop targeted SOP training materials; (harder) - finer changes in
abundance; finer changes in composition and age; ???

What Objectives are the decision makers hoping to achieve by using this
information?

Over what spatial scales is the information being used? e.g., at the
park, in the community, regional scale, flyway scale, nationally,
internationally For each spatial scale of interest, what are the
associated time scales of interest for detecting changes? E.g., annual,
w/in season, spring vs fall, etc.

From another report but quotes from Olsen et al. (1999): Olsen (et al.
1999) noted that ``Most of the thought that goes into a monitoring
program should occur at this preliminary planning stage. The objectives
guide, if not completely determine the scope of inference of the study
and the data collected, both of which are crucial for attaining the
stated objectives.'' Olsen goes on to say that a ``clear and concise
statement of monitoring objectives is essential to realize the necessary
compromises, select appropriate locations for inclusion in the study,
take relevant and meaningful measurements at these locations, and
perform analyses that will provide a basis for the conclusions necessary
for meeting the stated objectives.''

Fancy, Gross, and Carter (2009): The development of monitoring
objectives, which provide additional focus about the purpose or desired
outcome of the monitoring effort, was an iterative process that
sometimes required several years to refine. Early in the design process,
monitoring objectives were stated in more general terms, such as
``Determine trends in the incidence of disease and infestation in
selected plant communities and populations'', whereas the final
monitoring plan and protocols provided monitoring objectives that met
the test of being realistic, specific, and measurable (e.g., ``Estimate
trends in the proportion, severity, and survivorship of limber pine
trees infected with white pine blister rust at Craters of the Moon
National Monument''; Garrett et al. 2007).

Structure from Northeast Coastal and Barrier Network
\url{https://irma.nps.gov/DataStore/Reference/Profile/2195238}
Justification =\textgreater{} Vital Sign =\textgreater{} Monitoring Goal
=\textgreater{} Monitoring Questions =\textgreater{} Monitoring
Objectives =\textgreater{} Measures (eg pg 55 \& 56)

Clarify Objectives of the Decision Makers Among the decision maker
objectives identified in 1.b.ii, clarify the Fundamental Objective(s)
from the Means Objectives, then diagram the Objectives Hierarchy.
Identify the Attributes that need to be monitored to assess achievement
of each of the objectives. (`Concepts by Theory', Ford 2001) Identify
the Measurements selected to characterize each attribute. (`Concepts by
Data', Ford 2001) {[}WHAT{]} Identify any Threshold values associated w/
each Measurement

Q: compiling and assessing regional trends in waterbirds in SE AK? Asked
of Heather Renner (Ak Maritime), Robin Corcoran (Kodiak NWR) Answers:
Not really; regular discussion at PSG but no one has taken it on. Kathy
Kuletz works on statewide trends in at sea seabird observations; Piatt
GLBA work (\& Martin Renner); Robin C: ``There's always talk at the PSG
meeting about the North Pacific seabird colony database, or some
variation of that database for marine birds, but it never gets anywhere.
Basically there is no money for a database manager. I've been submitting
my data to Rob Kaler in Migratory Bird Management, but I think that's
just for storage until the money for a database manager materializes.
More recently Rob Suryan w/NOAA has been interested - for the North
Pacific Pelagic Seabird Database. You might want to contact him, he's
leading the Gulf Watch Alaska Long-Term Ecosystem Monitoring Program,
and maybe interested in this work.''

NPS/Audubon: Wu et al. 2018. Projected avifaunal responses to climate
change across the U.S. NPS. PLOS ONE. Appendices include projected
park-specific turnover, colonization, and extirpations projections -
potential to test with this survey. Would require discussions w/ CCRP \&
ARO - ideally as a coordinated effort with, say, AK Audubon and ACF(?)
or LCCs as tracking metric.\\
Appendices include ``Potential management goals and activities for
parks, organized by trend group.''

Melanie AK Audubon: IBA requires demonstrating higher than `usual'
densities (relative to surrounding area).

Nicole Michel \& Brooke Batemen Audubon National Science Team: of
interest and possible use in national scale quantitative analyses or
climate species distribution models. ``Scaling up and integrating
datasets across regions, including data from different goals, is
something that we're starting to do in other regions. ``

Occurrence, Richness Phenology (migration, breeding) Abundance - Even
without the structure needed to estimate detection probability, if the
same protocol is used annually this data could be used to produce
long-term trend indices for the study area. (competing staff priorities
- peak (?) usually aligns with hooligan run. Possibly reduce summer
frequency?)

Diagram the Conceptual Model of the system E.g., create an influence
diagram, being sure to denote the Fundamental \& Means Objective(s)
Factors influencing their achievement (aka `system drivers') Associated
management or policy decisions and/or potential actions. Include a
written summary of the literature upon which the conceptual model is
based.

Summarize the management or policy decisions the monitoring is to inform
(if any) based on 3.c.

\subsection{Phase 2 - Monitoring \& Analysis
Design}\label{phase-2---monitoring-analysis-design}

HOW monitor? Clarify the various design choices.

Clarify rationale for monitoring versus some other means of addressing
the information objectives (e.g., literature review, controlled
experiment, etc.), and type of monitoring being undertaken (status \&
trends or surveillance, threshold, effectiveness, or adaptive
management).

Questions: Why? What are information goals \& objectives? What sort of
observations would trigger action? What actions?

\subsubsection{Survey Design}\label{survey-design}

\emph{Target Frame} - space, time, scale, target populations
\emph{Spatial} Upper Lynn Canal visible from coastline from Skagway to
Dyea, including Skagway \& Taiya river mouths and Nahku Bay (between
Skagway and Dyea Point)

\begin{verbatim}
#> [1] "/home/mward2/klgo_bio_monitoring"
\end{verbatim}

Species: Targets: loons, grebes, cormorants, herons, ducks, geese,
plovers, sandpipers, gull, alcids, {[}kingfishers{]} Questions: clarify
role of kingfishers.

Incidental: 48 other spp Questions: are these all migratory? Any
year-round residents? Any of special concern? Any especially problematic
for detection on water or identification of spp?

(Hahr et al. \protect\hyperlink{ref-Hahr2004}{2004}), pg 13: Areas of
coastline that were visible from public access points between Skagway
and the Taiya River delta were divided into eight census units and
surveyed once per week in late April and May and bi-weekly during late
summer and fall (August -- mid September) (Figure 5). Surveys were
conducted as weather and scheduling permitted and dates/time were not
selected at random. Standardized observation points were established for
each census unit to ensure the most complete coverage of the study area,
and to minimize observer bias. Observers counted and identified all
birds visible within each unit and documented birds opportunistically
observed while moving between units. Counts were made by scanning the
census units with Bushnell Legend 10 x 42 binoculars and a Bausch \&
Lomb 20-60 x 70 spotting scope until the observer was able to quantify
all visible birds in the unit. Surveys were only conducted under optimal
weather conditions; however, occasionally conditions deteriorated during
the course of a survey making species identification difficult
(especially in the case of low clouds, fog and wind). Data were recorded
on standardized data forms developed after Collins et al. (2001)
(Appendix B). ArcView shapefiles of the observation points and census
units were created for use in the park's GIS. All data, original data
forms, and field notes were archived at KLGO.

\emph{Temporal} ??how define `season'??

Questions: what is cue to start of survey season? \emph{Madeleine later}
Diagnostic analysis =\textgreater{} check dates \& unit of first
non-zero counts each year for each spp (relative to dates of initiating
surveys)

\emph{Sample Frame} Spatial: same coverage as Target Frame. Area split
into eight survey areas (see Hahr et al.
\protect\hyperlink{ref-Hahr2004}{2004}). Questions: Access to all units?
Ease of observation of different units? Ease of distinguishing
boundaries? \emph{Madeleine later} Diagnostic analysis =\textgreater{}
value of all units - only some spp in unit X?; ``unit by species
occurrence'' relationships?

\emph{Sample Selection process} - any strata or clusters? Two-stage
processes? Probabilistic selection? {[}get in to enumerative vs
analytical distinction of Deming?{]} Currently survey all of frame
(`census' of units); no spatial sampling uncertainty. All spatial
uncertainty comes from Observation Process and detection issues.
Temporal sampling uncertainty remains.

Questions: feasible to reduce \# of units? See Qs above about `unique
contributions' of each unit. Logistic considerations? What effort level
is feasible? Diagnostic analysis =\textgreater{} any `representative'
units? How would that be defined?

\emph{Survey timing} Per season -

Questions: how define when to begin surveys? Zeros? Triggering events?
Potential confounding factors? (changes in phenology, ClCh); important
objective or confounder? Design options for addressing? Diagnostic
analysis=\textgreater{} Signs of confounding?

W/in season - (e.g., LEAU St.~George); potential refinement? Temporal:
(migration) 1 per wk, Apr 27 - May 20; (breeding) 1 per 2 wks, Jun 6 -
Jul 18; (migration) 1 per wk, Aug 1 - Sept 28. 2016: 17 surveys total
Questions: How define beginning \& end of breeding period? Info goals
from migration periods? From breeding periods? Design options to reduce
frequency? Diagnostic analysis=\textgreater{} .assessment of impact on
priority info goals of different frequencies? .spp specific occurrency
by survey date across years - capturing onset and end of migration?
.other assessments of survey timing relative to phenology of spp .mine
FB, Eagle Chat, others for date of first occurrences (see 2016 report);
Elaine. Requires clarifying information goals of survey vrs `first
occurrence'. E.g., not tail behavior.

Per survey route sequencing and tidal influence; SOP: 1,2,3,4,5,6,8,7 or
reverse, depending on tide timing (determines when do unit 7).

Questions: design options for confounding factors? Duration across eight
units: feasible to pull off? issue of bird movement between units?
Anyway to bound potential bias? Design options?

Monitoring Design Membership Design (panels, etc.) Single Panel

Revisitation Design W/in Year; varying effort levels (strata in time
across seasons) Between Years Always Repeat (pure panel), every year
Questions: feasbility of not doing every year? What are information
goals and triggers?

Response Design Attributes \& Measurements Focal targets vrs all species
Priority Spp for region? PSG? At Risk Spp (CCRP?) Audubon AK spp of
concern? See Audubon Climate Report. Relevant spp (intersection with spp
list)

\subsubsection{Observation Process}\label{observation-process}

Question: How determine boundaries? Visual markers. How handle movement
of birds into and out of unit? Check SOPs but basically `until things
stop'. Problem is that that is an accumulating approach that will bias
high areas with lots of productive habitat and/or well timed relative to
changing tidal conditions and sequencing of survey. How handle tendency
to record everything of note: e.g., recording observations in
neighboring units when really supposed to be focused on this unit. E.g.,
variation in time observing and effort observing across units \&
sampling events. This fails to track actual observation activity (time)
devoted to each unit as well as preferential bias to overestimate use in
areas. Think about Forsell \& Zweifelhofer? Chat w/ Emily \& Mark Otto
\& Andy? \#\#\#\# Covariates - captured \& not captured. Value, use in
analyses; more reliable alternative data sources; alignment of spatial
scales of interest and time scales of interest.

Lack of control over time at each site; portion spent traveling between
observation points, variation in number of observation points for each
site.

Variation in survey effort across sites, across per unit area, across
surveys (days), etc.

Major weather conditions (Northerlies, Southerlies) - simplify
conditions to these Unit 1 - recording and impact of presence \& \# of
cruise ships in dock, helicopter tours, etc., etc., other human
disturbances.

Distinguish value of covariates for act of (i) presence/absence
assessment, (ii) abundance count, (iii) composition \& breeding state.

Check SOP to clarify categorization of conditions.

Temp and wind greatly burden data collection process (slowing down
greatly). Variation in observer \#, training, and skill across surveys,
across years

\emph{Madeleine early} Diagnostic analyses =\textgreater{} .distribution
of time duration at each site by date and year .correlation across
units? .average surface area per unit .something about sequencing of
observation sites (e.g., confounding with tidal state)? .distribution of
observers by survey date by year .distribution of observer experience(?)
by survey date by year?

\paragraph{Detection issues associated with survey
conditions}\label{detection-issues-associated-with-survey-conditions}

.species specific counts w/in season by weather conditions and wave
conditions .species occurrence by wave conditions \& weather conditions
.species age and composition by wave conditions \& weather conditions
.total \# species by weather conditions, wave conditions; internally
consistent contrasts to demonstrate value, or lack thereof, of current
covariates. .richness by major weather conditions (Northerlies,
Southerlies, etc.) and for each spp .distribution of weather conditions
.distribution of survey time by weather conditions .diagnostic to ID key
survey times to ensure high quality observers? .consistency in `breeding
periods' across years (as captured in data) .quick diagnostics to ID
covariates seldom recorded, not consistently recorded, consistently
valued (e.g., no basis for contrast), etc.

Value of site-specific covariate recordings vs available `local' weather
(Skagway weather observations from airport - wind direction \& speed,
temp, etc.)

Ppt Codes: 0 - none, 1 - fog, 2 - drizzle, 3 - showers, 4 - light rain,
5 - moderate rain (steady), 6 - heavy rain, 7 - sleet, 8 - light snow, 9
- mod snow, 10 - heavy snow Beaufort scale codes: 0 - \textless{} 1 mph
(air calm); 1 - 1-3 mph (smoke drift show direction of wind); 2 - 4-7
mph wind felt on face, leaves rustle; 3 - 8-12 mph leaves \& small twigs
in constant motion; 4 - 13-18 raises dust, loose paper, moves small
branches; 5 - 19-24 small trees in leaf sway, crested wavelets form on
inland waters; Wave height: 0 - calm; 1 - scaly ripples, no foam crests;
2 - small wavelets, crests glassy, no breaking; 3 - large wavelets,
crests begin to break, scattered whitecaps; 4 - small waves, becoming
longer, numerous whitecaps; 5 - moderate waves, many whitecaps, some
spray SPP \# Age (Adult, imm, juv, mixed, unknown) Composition - M, F,
Fe/juve, Mixed, Unknown) Breeding Status - pairs, etcc; 24 potential
codings see SOP \#3.

Recommendations =\textgreater{} Clearly record standardized observation
locations on map and include in SOP to ensure consistency across
observers and personnel. Issue of veg growth reducing visible portion of
unit from each viewpoint - long-term confounding factor, etc. Value of
covariates, recommendations on what to drop, what to include Observation
quality notes to support assessment of extreme observations, etc.

Detection issues (seeing and recording available birds) Design options
for improving detection \& measurement Hmm? Lit? Emily, Aaron C., Kathi,
Mark Otto, Andy Royle? Feasible to do some sort of double observer
process - coordinate immediately after survey? Only conduct
intermittently per season rather than every time? E.g., like
sightability functions? Major drivers of sightability? Capturing
uncertainty on spp id Capturing uncertainty on spp count UAVs? Observer
training \& behavior classification Training materials from Kodiak or
Homer or Cordova or Haines or Juneau or Yakutat or\ldots{}?

\paragraph{Species identification
issues}\label{species-identification-issues}

Priority ID concerns: Commonly encountered species that can be difficult
Work with Elaine \& Heather \& Laura to develop list of commonly
confused species based on those commonly encountered (database or
Skagway Bird Club sightings) E.g., scoters, scaup, murrelets, etc.)
Target Spp of interest Training materials or other improvements Db error
checking? Other? Other expected sources of bias Species ID Behavior
classification Changes in equipment Summary Metrics Confounding factors
\& Covariates

\subsubsection{Analysis Design}\label{analysis-design}

Priority Summaries of interest (tie to Conceptual Model and Information
Objectives) Associated analyses:

Number of species sighted per year: 2003-2016 (source 2 found 2003-2009)
.Waterbird and non-waterbird species .Positive association between
species richness and year of survey, when excluding year 2014, which
started later than usual. R-squared = .54. .Species richness within
year: 2016 .Individual and max. numbers of observed counts for each
species for year 2016, with first and last sightings noted. .Peak
abundance/species richness identified with observed species counts and
individual sightings across surveys during 2016 season .Peak abundance
(May 4) observed to be in tandem with spring euchalon (April 21-)
.Species richness peaked May 4-11

Observer bias: variation in number of surveys and length of monitoring
season .2010: No seasonal technician; only sporadic surveys .2014:
Surveys begun 1 month late (May 22) .Across years: observer turnover
rate, varying experience

Number of surveys across census units per year (2003-2009) \&
distribution of survey dates

Percent of expected waterbird and breeding landbird species (174)
confirmed (158; 91\%)

Diversity .Simpson's Diversity Index (1/D) (2003-2009): heavily
sensitive to species richness .Q statistic (2003-2009) .Simpson evenness
measure: (1/D)/\# species, where values approaching 1 signal more equal
distribution (not sensitive to species richness)

Comparisons across census units: .Unit characteristics .Area of unit
(square km) .Shoreline habitat (meters) .Waterbird characteristics:
.Average number of waterbirds per survey per square km .Diversity
measures Analysis across dates each year: .Number of species; number of
birds .Diversity measures

Reporting schedule Questions: Who does the analyses? Interest in setting
up as Reproducible Research draft report? General delay in reporting?
When normally completed?

Secondary Summaries of interest Associated analyses Reporting schedule

Survey effort levels / Change detection targets / Power to detect
changes Reiterate targets for `local' inference vrs contribution to
regional trends assessments

1 Biotech conducts just this \& amphi monitoring.

\subsubsection{Data Management \& Reporting Structures / Software /
Workflows}\label{data-management-reporting-structures-software-workflows}

Data Models Design Requirements / User reporting needs

Elaine: Can put stuff in but can't get it out of existing ACCESS db.
Problematic. Bottleneck for analyses and summaries. Ebird - boundaries
don't include Skagway burough. Rather, formerly unincorporated area
including Skagway, Haines, and Angoon - too broad, not useful.

The current database is written in Microsoft Access version XX for
32-bit operating systems. Software and machine operating systems have
advanced enough for this to cause unnecessary burdens for accessing and
importing the data for analysis and reporting in common statistical
analysis environments (e.g., R). Recommend the Park request the ARO GIS
shop to convert the database to an uptodate software.

\emph{JOEL: Talk to Angie about what this would entail.}

\emph{Recommendations} work with GIS shop to clean up db, improve
reporting \& query support; shiny app; reproducible research scripts to
generate automatic reports? Shiny app to interface?

\section{Discussion}\label{discussion}

\section{Conclusion}\label{conclusion}

\section{Acknowledgements}\label{acknowledgements}

\subparagraph{pagebreak}\label{pagebreak}

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Anonymous2009}{}
Anonymous, (KLGO NHP). 2009. ``Waterbird Monitoring Protocol for
Klondike Gold Rush National Historical Park , Skagway , Alaska Standard
Operating Procedure ( SOP ) \# 3 Conducting the Waterbird Survey
Transect Revision History Log :'' April 9, 2009.

\hypertarget{ref-Fancy2009}{}
Fancy, Steven G., J. E. Gross, and S. L. Carter. 2009. ``Monitoring the
condition of natural resources in US national parks.''
\emph{Environmental Monitoring and Assessment} 151 (1-4): 161--74.
doi:\href{https://doi.org/10.1007/s10661-008-0257-y}{10.1007/s10661-008-0257-y}.

\hypertarget{ref-Hahr2004}{}
Hahr, Meg, Todd W Trapp, Meg Hahr, and Todd W Trapp. 2004. ``Waterbird
and Breeding Landbird Inventories in Klondike Gold Rush National
Historical Park Waterbird and Breeding Landbird Inventories in Klondike
Gold Rush National Historical Park Final Report,'' no. May.

\hypertarget{ref-Mitchell2018}{}
Mitchell, Brian, Alice Chung-MacCoubrey, Jim Comiskey, Lisa Garrett,
Maggie MacCluskie, Bill Moore, Tom Philippie, Geoff Sanders, and John
Paul Schmit. 2018. ``Inventory and Monitoring Division Protocol Review
Guidance.'' Fort Collins, CO.

\hypertarget{ref-NABCI2000}{}
(NABCI), North American Bird Conservation Initiative. 2000. ``North
American Bird Conservation Initiative Bird Conservation Region Map and
Descriptions.'' Arlington, VA.
\url{http://nabci-us.org/resources/bird-conservation-regions/}.

\hypertarget{ref-Reynolds2012}{}
Reynolds, Joel H. 2012. ``An overview of statistical considerations in
long-term monitoring.'' In \emph{Design and Analysis of Long-Term
Ecological Monitoring Studies}, 23--53.
doi:\href{https://doi.org/10.1017/CBO9781139022422.005}{10.1017/CBO9781139022422.005}.

\hypertarget{ref-Reynolds2016}{}
Reynolds, Joel H., Melinda G. Knutson, Ken B. Newman, Emily D.
Silverman, and William L. Thompson. 2016. ``A road map for designing and
implementing a biological monitoring program.'' \emph{Environmental
Monitoring and Assessment} 188 (7). Environmental Monitoring;
Assessment: 1--25.
doi:\href{https://doi.org/10.1007/s10661-016-5397-x}{10.1007/s10661-016-5397-x}.

\hypertarget{ref-St.Pierre2015}{}
St. Pierre, Mallory (KLGO NHP). 2015. ``2014 Bird Surveys at Klondike
Gold Rush National Historical Park.'' Fort Collins: U.S. Department of
Interior, National Park Service, NRSS.

\hypertarget{ref-Surdyk2017}{}
Surdyk, Shelby L. (KLGO NHP). 2017. ``Bird Surveys at Klondike Gold Rush
NHP: 2015 summary.''

\hypertarget{ref-Surdyk2018}{}
Surdyk, Shelby L. (KLGO NHP), and S. A. (KLGO NHP) Evans. 2018. ``Bird
Surveys at Klondike Gold Rush NHP Bird Surveys at Klondike Gold Rush
NHP: 2016 summary.'' Fort Collins: National Park Service.

\hypertarget{ref-Wu2018}{}
Wu, J X, C B Wilsey, L Taylor, and G W Schuurman. 2018. ``Projected
avifaunal responses to climate change across the U.S. National Park
System.'' \emph{PLoS ONE}.
doi:\href{https://doi.org/10.1371/journal.pone.0190557}{10.1371/journal.pone.0190557}.

\subparagraph{pagebreak}\label{pagebreak-1}

\subsubsection{Colophon}\label{colophon}

This report was generated on 2018-12-10 13:33:40 using the following
computational environment and dependencies:

\begin{verbatim}
#>  setting  value                       
#>  version  R version 3.5.1 (2018-07-02)
#>  system   x86_64, linux-gnu           
#>  ui       X11                         
#>  language (EN)                        
#>  collate  en_US.UTF-8                 
#>  tz       EST5EDT                     
#>  date     2018-12-10                  
#> 
#>  package    * version date       source        
#>  assertthat   0.2.0   2017-04-11 CRAN (R 3.3.1)
#>  backports    1.1.2   2017-12-13 CRAN (R 3.5.0)
#>  base       * 3.5.1   2018-09-10 local         
#>  bindr        0.1.1   2018-03-13 CRAN (R 3.5.0)
#>  bindrcpp   * 0.2.2   2018-03-29 CRAN (R 3.5.0)
#>  colorspace   1.3-2   2016-12-14 CRAN (R 3.5.0)
#>  compiler     3.5.1   2018-09-10 local         
#>  crayon       1.3.4   2017-09-16 CRAN (R 3.3.1)
#>  datasets   * 3.5.1   2018-09-10 local         
#>  DBI          1.0.0   2018-05-02 CRAN (R 3.5.0)
#>  dbplyr     * 1.2.2   2018-07-25 CRAN (R 3.5.0)
#>  devtools     1.13.6  2018-06-27 CRAN (R 3.5.0)
#>  digest       0.6.17  2018-09-12 CRAN (R 3.5.0)
#>  dplyr      * 0.7.6   2018-06-29 CRAN (R 3.5.0)
#>  evaluate     0.11    2018-07-17 CRAN (R 3.5.0)
#>  farver       1.1.0   2018-11-20 CRAN (R 3.5.1)
#>  ggforce    * 0.1.3   2018-07-07 CRAN (R 3.5.1)
#>  ggplot2    * 3.0.0   2018-07-03 CRAN (R 3.5.0)
#>  ggthemes   * 4.0.1   2018-08-24 CRAN (R 3.5.1)
#>  glue         1.3.0   2018-07-17 CRAN (R 3.5.0)
#>  graphics   * 3.5.1   2018-09-10 local         
#>  grDevices  * 3.5.1   2018-09-10 local         
#>  grid         3.5.1   2018-09-10 local         
#>  gtable       0.2.0   2016-02-26 CRAN (R 3.5.0)
#>  hms          0.4.2   2018-03-10 CRAN (R 3.5.0)
#>  htmltools    0.3.6   2017-04-28 CRAN (R 3.5.0)
#>  knitr        1.20    2018-02-20 CRAN (R 3.5.0)
#>  lazyeval     0.2.1   2017-10-29 CRAN (R 3.5.0)
#>  lubridate  * 1.7.4   2018-04-11 CRAN (R 3.5.0)
#>  magrittr     1.5     2014-11-22 CRAN (R 3.1.2)
#>  MASS         7.3-50  2018-04-30 CRAN (R 3.5.1)
#>  memoise      1.1.0   2017-04-21 CRAN (R 3.4.3)
#>  methods    * 3.5.1   2018-09-10 local         
#>  munsell      0.5.0   2018-06-12 CRAN (R 3.5.0)
#>  pillar       1.3.0   2018-07-14 CRAN (R 3.5.0)
#>  pkgconfig    2.0.2   2018-08-16 CRAN (R 3.5.0)
#>  plyr         1.8.4   2016-06-08 CRAN (R 3.5.0)
#>  purrr        0.2.5   2018-05-29 CRAN (R 3.5.0)
#>  R6           2.2.2   2017-06-17 CRAN (R 3.3.1)
#>  Rcpp         0.12.18 2018-07-23 CRAN (R 3.5.0)
#>  readr      * 1.1.1   2017-05-16 CRAN (R 3.5.0)
#>  rlang        0.2.2   2018-08-16 CRAN (R 3.5.0)
#>  rmarkdown    1.10    2018-06-11 CRAN (R 3.5.0)
#>  rprojroot    1.3-2   2018-01-03 CRAN (R 3.5.0)
#>  scales       1.0.0   2018-08-09 CRAN (R 3.5.0)
#>  stats      * 3.5.1   2018-09-10 local         
#>  stringi      1.2.4   2018-07-20 CRAN (R 3.5.0)
#>  stringr    * 1.3.1   2018-05-10 CRAN (R 3.5.0)
#>  tibble       1.4.2   2018-01-22 CRAN (R 3.5.0)
#>  tidyr      * 0.8.1   2018-05-18 CRAN (R 3.5.0)
#>  tidyselect   0.2.4   2018-02-26 CRAN (R 3.5.0)
#>  tools        3.5.1   2018-09-10 local         
#>  tweenr       1.0.0   2018-09-27 CRAN (R 3.5.1)
#>  units        0.6-0   2018-06-09 CRAN (R 3.5.0)
#>  utils      * 3.5.1   2018-09-10 local         
#>  withr        2.1.2   2018-03-15 CRAN (R 3.5.0)
#>  yaml         2.2.0   2018-07-25 CRAN (R 3.5.0)
\end{verbatim}

The current Git commit details are:

\end{document}